The watershed transform in ITK - discussion and new developments

### 2006-04-23 article/Article.tex
article continuated
darcs-hash:20060423203001-2fc9d-8c629973fe2020663d13cf8f6a114ef38d1395f6.gz

The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
the pixels: $already added to the queue$ or $not yet added to the queue$.
This task is often done by using negative values in the output image
for the pixels not yet added to the queue. In our implementation, we have
chosen to use a binary image to store this, while labels always take values
above 0 and watershed lines are marked with value 0. The negative values
option have several drawbacks:
\begin{itemize}
  \item it requires to use a different image type than the label type defined
by the user in case the user have chosen an unsigned pixel type.
  \item it requires to use one of the value which can be possibly aready used
by a label in the marker image, in case of a signed type.
\end{itemize}
Those problems can be avoided\footnote{It's not true for the unsigned long
pixel type} by using a type which contains a larger range of value than the
type choosed by the user, but then, we must copy the data to an image of
the type chosen by the user at the end of the algorithm.

@@

@@

\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
without marker. It encapsulate several filters:
\begin{itemize}
  \item \verb$HMinimaImageFilter$
  \item \verb$RegionalMinimaImageFilter$
  \item \verb$ConnectedComponentImageFilter$
  \item \verb$MorphologicalWatershedFromMarkersImageFilter$
\end{itemize}
While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
to avoid the classical watershed's over segmentation by allowing the user
to select the dynamic of the minima.

@@

>>>>

The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
the pixels: {\em already added to the queue} or {\em not yet added to the queue}.
This task is often done by using negative values in the output image
for the pixels not yet added to the queue. In our implementation, we have
chosen to use a binary image to store this, while labels always take values
above 0 and watershed lines are marked with value 0. The negative values
option have several drawbacks:
\begin{itemize}
  \item it requires to use a different image type than the label type defined
by the user in case the user have chosen an unsigned pixel type.
  \item it requires to use one of the value which can be possibly aready used
by a label in the marker image, in case of a signed type.
\end{itemize}
Those problems can be avoided\footnote{It's not true for the unsigned long
pixel type} by using a type which contains a larger range of value than the
type choosed by the user, but then, we must copy the data to an image of
the type chosen by the user at the end of the algorithm.

@@

The connectivity is defined using the usual ITK way: \verb$SetFullConnected(false)$
(4-connected in 2D, 6-connected in 3D) or \verb$SetFullConnected(true)$ (8-connected
in 2D, 26-connected in 3D)\footnote{One should regret to not have the 18-connected
connectivity in 3D}.

@@

\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
without marker. It encapsulate several filters:
\begin{itemize}
  \item \verb$HMinimaImageFilter$
  \item \verb$RegionalMinimaImageFilter$
  \item \verb$ConnectedComponentImageFilter$
  \item \verb$MorphologicalWatershedFromMarkersImageFilter$
\end{itemize}
While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
to avoid the classical watershed's over segmentation by allowing the user
to select the dynamic of the minima. The minimum dynamic of minima can be
selected with \verb$SetThreshold()$.

@@

\begin{figure}[htbp]
\begin{center}
\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-rgb}}
\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-rgb}}
\subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-rgb}}
\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=4]{threshold30-rgb}}
\subfigure[Minima threshold = 50]{\includegraphics[scale=4]{threshold50-rgb}}
\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markersMf-rgb}}
\caption{Different results of watershed on an articial image. The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}. From $(b)$ to $(f)$, an example of reduction of the classical waterhsed oversegmentation by minima dynamic filtering. $(g)$ and $(h)$ show a marker defined by a user, and the result over the input image. All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\subfigure[]{\includegraphics[scale=10]{pass-values}}
\subfigure[]{\includegraphics[scale=10]{pass-values-mf-rgb}}
\subfigure[]{\includegraphics[scale=10]{pass-values-mF-rgb}}
\subfigure[]{\includegraphics[scale=10]{pass-values-Mf-rgb}}
\subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
\subfigure[]{\includegraphics[scale=10]{pass-values}}

\subfigure[]{\includegraphics[scale=4]{plateau}}
\subfigure[]{\includegraphics[scale=4]{plateau-mf-rgb}}
\subfigure[]{\includegraphics[scale=4]{plateau-mF-rgb}}
\subfigure[]{\includegraphics[scale=4]{plateau-Mf-rgb}}
\subfigure[]{\includegraphics[scale=4]{plateau-MF-rgb}}
\subfigure[]{\includegraphics[scale=4]{plateau}}

\subfigure[]{\includegraphics[scale=10]{thick-lines}}
\subfigure[]{\includegraphics[scale=10]{thick-lines-mf-rgb}}
\subfigure[]{\includegraphics[scale=10]{thick-lines-mF-rgb}}
\subfigure[]{\includegraphics[scale=10]{thick-lines-Mf-rgb}}
\subfigure[]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
\subfigure[]{\includegraphics[scale=10]{thick-lines}}

\subfigure[]{\includegraphics[scale=6]{button-hole}}
\subfigure[]{\includegraphics[scale=6]{button-hole-mf-rgb}}
\subfigure[]{\includegraphics[scale=6]{button-hole-mF-rgb}}
\subfigure[]{\includegraphics[scale=6]{button-hole-Mf-rgb}}
\subfigure[]{\includegraphics[scale=6]{button-hole-MF-rgb}}
\subfigure[]{\includegraphics[scale=6]{button-hole}}

\caption{}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\subfigure[]{\includegraphics[scale=0.7]{cthead1}}
\subfigure[]{\includegraphics[scale=0.7]{cthead1-markers}}
\subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
\subfigure[]{\includegraphics[scale=0.7]{cthead1mF-rgb}}
\subfigure[]{\includegraphics[scale=0.7]{cthead1Mf-rgb}}
\subfigure[]{\includegraphics[scale=0.7]{cthead1MF-rgb}}
\caption{}
\end{center}
\end{figure}
